---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE,
  error = FALSE,
  fig.width = 8,
  fig.height = 8)
```

# async

> Asynchronous HTTP

<!-- badges: start -->
![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)
[![Codecov test coverage](https://codecov.io/gh/r-lib/async/graph/badge.svg)](https://app.codecov.io/gh/r-lib/async)
[![R-CMD-check](https://github.com/r-lib/async/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-lib/async/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The async package brings asynchronous (async) computation and I/O to R.
It uses an event loop to schedule asynchronous functions that
report their results via deferred values. Deferred values can be chained
together for complex async computation, and they are evaluated lazily,
at synchronisation points.

## Features

- A `deferred` class which is the basic building block for async
  computation.
- Timers, HTTP queries, generic external processes, external R processes.
- A worker pool for calling R functions in the background.
- Operations to combine deferred values: `$then()`, `$when_all()`,
  `when_any()`, `when_some()`, `$finally()` and the `$catch()`
  operation to handle errors.
- A strong ownership model: each deferred value has exactly one child
  that receives the result (or error) of its async computation.
- Synchronisation (sync) points to embed async code into synchronous code.
  Sync points run their own event loop, and they can be stacked.
- Lazy evaluation of deferred values. Only the deferred values that are
  needed to calculate the result of the async phase are evaluated.
- Auto-cancellation. Deferred values that are not needed for the final
  result(s) of the async phase are automatically cancelled, as soon as
  possible.
- Manual cancellation is also possible.
- Progress bar support. Deferred values can report their progress via a
  callback function.
- Additional helper functions for working with deferred values, e.g.
  `async_map()`, `async_detect()`, `async_filter()`, etc.

## Supported async I/O and computation

We support the following async primitives:
- Timers: `delay()`.
- HTTP queries: `http_get()`, `http_head()`.
- External processes: `run_process()`.
- External R processes: `run_r_process()`.
- A workers pool of processes to evaluate R code: `call_function()`.

## Introduction

The async package brings asychronous I/O and computation to R. It implements
asynchronous HTTP requests, timers, subprocesses and an R worker pool.

## Deferred Values

Asynchronous computation is represented by deferred values. In async
a deferred value is an [R6](https://github.com/wch/R6) object, so it has
reference semantics. In some other programming languages deferred values
are called futures or promises.

You can think of a deferred value as a placeholder for a (non-deferred)
value that is not yet known. When the actual value of a deferred is
computed, we say that the deferred is resolved.

Typically a deferred value is created by requesting asynchronous I/O, like
an HTTP GET request with the `http_get()` function.

## Built-in Deferred Value Constructors

The async package has built-in async functions that create deferred values:

- `delay()` creates a timer that expires after the specified time.
- `http_get()` and `http_head()` perform HTTP requests, asynchronously.
- `async_constant()` creates a simple deferred that represents the supplied
  value.
- `run_process()` runs an external process using processx and returns
  its exit code, standard output and error, asynchronously.
- `run_r_process()` runs an external R process, and calls the specified
  R function in this process. It returns its exit status, standard output,
  standard error, and the return value of the R function call,
  asynchronously.
- `call_function()` uses a worker pool of persistent external R processes
  to call R functions. It returns the return value of the function, and
  the standard output and error of the process, asynchronously.

## Deferred chains

While the actual value of a deferred cannot be queried directly,
we can still operate on it, by declaring code that is to be executed,
once the value will be known:

```{r}
library(async)
http_status <- function(url) {
  def <- http_get(url)$
	then(function(response) response$status_code)
}
synchronise(http_status("https://httpbin.org"))
```

`http_status()` is an async function that returns the status code of a GET
HTTP request to the specified URL. It works by creating an async GET
request and then specifying that once the HTTP response in, the status
code should be accepted from it. The deferred value returned by
`http_status()` resolves to the status code of the URL.

`then()` returns another deferred, which also has a `then()` method,
so it is chainable.

`$then()` is the simplest combination operator on deferred values.

`$when_all()` is similar, but it creates a deferred value that resolves
once all deferred values passed to `$when_all()` are computed.

`$when_any()` creates a deferred value that resolves as soon as one of
its arguments successfully resolves. `when_some()` is its generalization
that requires the computation of a given number of deferred values.

## Synchronization barriers

async allows embedding asynchronous computation in synchronous code.
The execution of such a program has a sync phase and async phases. When the
program starts, it is in the sync phase. In the sync phase you cannot
create deferred values. (But you can still define (async) functions, that
will create deferred values when called.)

To enter into an async phase, call `synchronise()` on an expression that
evaluates to a deferred value. The async phase will last until this
deferred value is computed or an error is thrown (and the error reaches
`synchronise()`).

`synchronise()` creates an event loop, which manages the computation of
the deferred values in this particular async phase.

Async phases can be embedded into each other. I.e. a program may call
`synchronise()` while in the async phase. The outer async phase's event
loop then stops until the inner async phase terminates. Deferred values
cannot be passed through a `synchronise()` barrier, to anoter (sync or
async phase). Should this happen, an error is reported on the first
operation on the leaked deferred value.

In a typical application, a function is implemented asynchronously, and
then used synchronously by the interactive user, or another piece of
synchronous code, via `synchronise()` calls. The following example makes
three HTTP requests in parallel:

```{r}
http_status3 <- function() {
  http_status <- function(url) {
    http_get(url)$then(function(response) response$status_code)
  }
  r1 <- http_status("https://httpbin.org/status/403")
  r2 <- http_status("https://httpbin.org/status/404")
  r3 <- http_status("https://httpbin.org/status/200")
  when_all(r1, r2, r3)
}
synchronise(http_status3())
```

## Error handling

There are two ways to handle errors in asynchronous code.
One is the `$catch()` operation that can be called on a deferred value.
`$catch()` has similar syntax as `tryCatch()`. It can be used to catch
errors during the computation of the deferred value, including computation
in its ancestors (except for the errors the ancestors already handle).

```{r}
response_time <- async(function(url) {
  http_head(url)$
    then(http_stop_for_status)$
    then(function(x) setNames(x[["times"]][["total"]], url))$
    catch(error = function(...) setNames(Inf, url))
})
synchronise(response_time("https://google.com"))
synchronise(response_time("https://httpbin.org/status/401"))
```

Errors can also be handled synchronously. If an error is not handled
asynchronously, then the deferred value will throw an error when itself
or one of its descendants is `synchronise()`-ed. This can be caught with
`tryCatch()`.

## Ownership and the Async DAG

When the `$then()` method of a deferred value is called to create another
deferred value:

```r
d2 <- d1$then(function(x) ...)
```

then we say that `d2` owns `d1`. We also say that `d2` is the child of `d1`,
and `d1` is the parent of `d2`. async has a strong ownership model, and it
only allows a single owner (i.e. a single child) for each deferred.
The parent-child relationships define a directed forest graph, a
collection of directed trees. (This is without shared deferred values,
see the manual.)

The strong ownership model does not allow calling `$then()` multiple times
on the same deferred value, i.e. the following generates an error:
```{r, error = TRUE}
do <- function() {
  d <- delay(1/100)
  d$then(function() print("foo"))
  d$then(function() print("bar"))
}
synchronise(do())
```

The `when_all()`, `when_any()` and `when_some()` operations set a single
deferred as the owner of multiple parents. For `when_all()` the child
node is resolved once all of its parents are resolved (or one throws an
error). `when_any()` resolves as soon as one of its parents resolves. If all
of its parents throw errors then `when_any()` throws as well.
`when_some()` is a generalization of `when_any()` and it resolves as
soon as the specified number of its parents resolve without error, or
if too many parents fail for `when_some()` to be successful.

When `synchronise()` is called on a deferred value, the DAG rooted there
is called the async DAG of the async phase. (This is usually a directed
tree, and in this README we do not deal with shared deferred values,
which would result more general DAGs.)

When the strict shared ownership model is too restrictive, certain
deferred values can be marked as shared, via the `$share()` method.
These can have multiple owners (children) and they are also not
auto-cancelled (see Auto-Cancellation later).

## Lazy Evaluation

async does not evaluate deferred values that are not part of the async
DAG of the async phase. These are clearly not needed to compute the
result of the async phase, so it would be a waste of resources working on
them. (It is also unclear how their errors should be handled.)

In the following example, `d1` and `d2` are created, but they are not
part of the async DAG, so they are never evaluated.

```{r}
do <- function() {
  d1 <- delay(1/100)$then(function() print("d1"))
  d2 <- d1$then(function() print("d2"))
  d3 <- delay(1/100)$then(function() print("d3"))
  d4 <- d3$then(function() print("d4"))
  d4
}
invisible(synchronise(do()))
```

## Auto-Cancellation

In an async phase, it might happen that parts of the async DAG are not
needed for the final result any more. E.g. if a parent of a `when_all()`
node throws an error, then the other parents don't have to be computed.
In this case the event loop of the phase automatically cancels these
deferred values. Similarly, if a single parent of a `when_any()` node is
resolved, the other parents can be cancelled.

In general, if a node of the async DAG is resolved, the whole directed
DAG, rooted at that node, can be cancelled (except for nodes that were
already resolved and nodes that have already failed).

Auto-cancellation is very convenient, as you can be sure that resources
are free as soon as they are not needed. Some practical examples:

* Making HTTP requests to many mirror web sites, to check their response
  time. As soon as the first reply is in, the rest of the HTTP requests
  are cancelled.
* In multi-process computation, as soon as one process fails, the rest are
  automatically cancelled. (Unless the failure is handled, of course.)

async also has another type of cancellation, when `synchronise()` is
interrupted externally, either by the user or some system error. In this
case all processes and resources that were created in the event loop,
are cancelled and freed.

Shared deferred values (see `$share()`) are not auto-cancelled when their
children are resolved or errored, but they are always cancelled at the
end of the async phase.

## Async Iterators

async provides some utilities that make it easier to deal with
collections of deferred values. E.g. `async_map()` applies an async function
to a list and returns a single deferred value for the whole result.
`async_detect()` finds a value in a list that satisfies an async predicate
function, etc.

The current iterators:
* `async_map()` applies an async function to all elements of a vector or
  list (collection).
* `async_detect()` finds an element of a collection that passed an async
  truth test.
* `async_every()` checks if every element of a collection satisfies an async
  predicate. `async_some()` checks if any element does that.
* `async_filter()` keeps elements that pass an async truth test.

## Async Control Flow

Control flow with deferred values can be challenging. Some helpers:
* `async_reflect()` creates an async function that always succeeds.
  This is useful if you want to apply it to a collection, and don't
  want to stop at the first error.
* `async_retry()` tries an async function a number of times.
  `async_retryable()` turns a regular function into a retryable one.
* `async_sequence()` chains two async functions. Calling their sequence is
  equivalent calling `then()` on them, but `async_sequence()` is easier to
  use programmatically.
* `async_until()` and `async_whilst()` let you call an async function
  repeatedly, until or while a (syncronous) condition holds.
* `async_timeout()` runs an async function with a timeout.

## Other Async Utilities

* `async_constant()` takes a value and creates and asynchronous function
  that returns that value.

## Examples

### Parallel API Queries

Query the crandb API, get the authors of the packages with the most
reverse dependencies.

```{r}
fromJSON <- function(x) jsonlite::fromJSON(x, simplifyVector = FALSE)
revdep_authors <- function() {
  get_author <- function(package) {
    url <- paste0("https://crandb.r-pkg.org/", package)
    http_get(url)$
      then(function(x) fromJSON(rawToChar(x$content)))$
      then(function(x) x$Author)
  }

  http_get("https://crandb.r-pkg.org/-/topdeps/devel")$
    then(function(x) fromJSON(rawToChar(x$content)))$
    then(function(x) names(unlist(x)))$
    then(function(x) async_map(x, get_author))
}
synchronise(revdep_authors())[1:3]
```

### Checking URLs

The following code returns the 2 URLs that respond with the shortest
response time.

```{r}
response_time <- async(function(url) {
  http_head(url)$
    then(http_stop_for_status)$
    then(function(x) setNames(x[["times"]][["total"]], url))$
    catch(error = function() setNames(Inf, url))
})

fastest_urls <- async(function(urls, n = 2) {
  reqs <- lapply(urls, response_time)
  when_some(n, .list = reqs)$
    then(function(x) sort(unlist(x)))
})

urls <- c("https://cran.rstudio.com", "https://cran.r-project.org",
          "https://www.stats.bris.ac.uk/R/", "https://cran.uib.no/")

synchronise(fastest_urls(urls))
```

See the package vignettes for more examples.

## License

MIT © [RStudio Inc](https://rstudio.com)
